---
title: "PRoMAn: A Research Production System"
author: "John Grady Heller"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: false
      smooth_scroll: true
    theme: flatly
    highlight: tango
vignette: >
  %\VignetteIndexEntry{PRoMan: A Complete Digital Research Production System}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
library(PRoMan)
```

PRoMan is a comprehensive digital research production system that transforms academic workflow through intelligent project management, automated analysis orchestration, and the revolutionary **Blitzschreiben** methodology. Built specifically for ADHD-friendly workflows, PRoMan provides structure, automation, and continuity for the entire research lifecycle.

## System Overview

PRoMan integrates six major components:

1.   **Project Management** - Standardized structure and intelligent file handling

2.  **Blitzschreiben Methodology** - Three-stage research production pipeline

3.  **Analysis Workflow Orchestration** - Master scripts and component-based analysis

4.  **Intelligent Activity Tracking** - Automated logging and session management

5.   **Publication Support** - Figure management and academic formatting

6.   **Collaboration Tools** - Onboarding packages and communication templates

---

# Part I: Core Project Management

## Installation and Setup

```{r install}
# Install from local development folder
devtools::install("path/to/PRoMan")

# Load PRoMan (automatically enables multi-project session tracking)
library(PRoMan)
```

## Basic Project Creation

```{r basic_project}
# Create a standard project structure
paths <- create_project("my_research_project")

# This creates:
# my_research_project/
# ├── data/
# │   ├── source/          # Raw data files
# │   └── cleaned/         # Processed data
# ├── r_programs/          # R scripts and analysis code
# ├── personal_notes/      # Documentation and notes
# │   ├── analyseblitz/    # Quick analysis notes
# │   └── figures/         # Figure registry and outputs
# ├── sources/             # Reference materials (PDFs, articles)
# ├── old/                 # Archived files
# └── analyse.qmd          # Main analysis document
```


## Automatic Project Detection

PRoMan's intelligent project detection means you never need to set working directories or remember paths:

```{r auto_detection}
# From ANY subfolder in your project, PRoMan finds the root
# Example: You open a script in project/r_programs/analysis.R

paths <- get_project_paths()  # Automatically detects project root
# Message: "Found project root: my_research_project"
#         "  (detected from: r_programs)"

# Works from any depth
# Even from project/data/source/subfolder/deep/
paths <- get_project_paths()  # Still finds the root!

# The detection looks for required folders: data/, personal_notes/, r_programs/

```

## Intelligent File Set Management

PRoMan excels at managing groups of related files - useful for multi-state, multi-year, or multi-condition datasets:

```{r file_sets}
# Find all files containing specific keywords
count_files <- select_files_by_keyword("counts")
# Returns: c("data_MD_counts.csv", "data_VA_counts.csv", ...)

# Multiple keywords (must contain ALL)
state_summary_files <- select_files_by_keyword(c("state", "summary"))

# Load entire file sets at once
state_data <- load_files_by_keyword("state")
# Returns named list: state_data$data_MD_state, state_data$data_VA_state, etc.

# Load from specific folder
clean_data <- load_files_by_keyword("processed", 
                                   folder_path = paths$data_cleaned)
```

## Creating File Sets

PRoMan provides multiple ways to create file sets:

```{r create_sets}
# Method 1: Specify files directly
create_file_set(c("data_2021.csv", "data_2022.csv"), 
                set_name = "annual")

# Method 2: Interactive selection
create_file_set("interactive", set_name = "baseline")
# Opens file browser for selection

# Method 3: Recent files (last 30 minutes)
create_file_set("recent:30", set_name = "just_processed")

# Method 4: Use keywords if exact names unknown
create_file_set(c("survey", "baseline"), set_name = "wave1")
# Finds all files matching BOTH keywords

# List detected file sets
list_file_sets()
# Shows organized groups like "State files", "Count files", etc.

# Combine existing sets with new identifier
add_to_sets(c("wave1", "wave2"), 
            new_identifier = "combined",
            archive_originals = TRUE)

```

## Automatic File Set Detection

```{r set_detection}
# PRoMan automatically detects file patterns
analyze_file_sets()  # Shows all detected sets in project

# Detection includes:
# - State files (ending in _XX state codes)
# - Time series (files with years)
# - Processing stages (raw, clean, final)
# - Common prefixes (automatically grouped)

# Get file sets programmatically
file_sets <- get_file_sets()
# Returns nested list by folder and set type
```

## Advanced Data Cleaning and Quality Assessment

```{r data_quality}
# Comprehensive data quality assessment
data_snapshot(my_data, save_summary = TRUE)

# Detect suspicious values (common missing data codes)
suspicious <- flag_suspicious_values(my_data, 
                                   suspicious_codes = c(-999, -99, 999))

# Analyze missing data patterns
missing_analysis <- missing_data_patterns(my_data, threshold = 5)

# Auto-detect ID columns
id_cols <- detect_id_columns(my_data)

# Standardize variable names to snake_case
clean_data <- standardize_var_names(my_data, naming_convention = "snake_case")

# Clean demographic variables automatically
clean_data <- clean_demographics(my_data, 
                                age_var = "participant_age",
                                gender_var = "gender", 
                                race_var = "ethnicity")

# Generate comprehensive data dictionary
data_dict <- generate_data_dictionary(clean_data, 
                                    filename = "data_dictionary.csv",
                                    include_stats = TRUE,
                                    guess_descriptions = TRUE)
```

## String Cleaning Utilities

```{r string_cleaning}
# Standardize messy strings for joining
messy_locations <- c(" maryland ", "VIRGINIA", "new york ", "  North Carolina  ")
clean_locations <- clean_strings(messy_locations)
# Result: c("MARYLAND", "VIRGINIA", "NEW_YORK", "NORTH_CAROLINA")

# Clean data frame columns
df <- clean_strings(df, "state_name", "state_clean")  # Create new column
df <- clean_strings(df, "location_var")               # Overwrite original
```

## File Management and Organization

```{r file_management}
# Archive old files with automatic dating
archive_files(c("old_analysis.R", "temp_data.csv"))

# Find files that haven't been touched recently
stale_files <- find_stale_files(days = 30)

# Check for large files taking up space
large_files <- find_large_files(min_size_mb = 50)

# Get detailed folder information
folder_summary <- get_folder_summary(sort_by = "date")

# Verify required files exist before analysis
check_required_files(c("main_data.csv", "lookup_table.csv"))

# Save data with automatic timestamps
save_with_timestamp(analysis_results, "final_model")
# Creates: final_model_20231201_143022.csv

# Compare datasets before and after processing
comparison <- compare_datasets(original_data, processed_data, 
                             id_column = "participant_id")
```

---

# Part II: The Blitzschreiben Methodology

The heart of PRoMan is the **Blitzschreiben** system - a three-stage research production methodology designed for rapid, systematic research execution.

## Blitzschreiben Overview

**Philosophy**: Move ideas from conception to publication through structured, time-bounded stages with clear decision points.

**Weekly Rhythm**:
- **Monday**: Stage I Triage + Stage II Prep
- **Tuesday**: Stage II Mock Analysis
- **Wednesday**: Stage II Real Analysis
- **Thursday**: Stage III Draftschreiben
- **Friday**: QA Review + Source Reading

**Pipeline Management**: Maintain 4-5 projects in Stage I, 3 in Stage II, 2 in Stage III

## Stage I: Idea Triage (12-15 minutes per idea)

Rapid evaluation to determine which ideas merit further development.

```{r stage1}
# Create new Blitzschreiben project
create_blitz_project("Social Media Impact on Academic Performance", "P4.1")

# Start timed triage session
start_triage_timer(minutes = 15, project_code = "P4.1")

# Check remaining time during triage
check_triage_time()

# Complete triage with decision
end_triage_timer(decision = "ready_to_share", 
                next_steps = "Find social media usage datasets")
```

**Stage I Card Template includes**:
- Research question formulation
- Data/dataset identification
- Sample description (N, demographics, key variables)
- Analysis plan (models, adjustments, visualizations)
- Background warrant (literature gap, importance)
- Current barriers and challenges
- **Decision**: Ready to share / Needs development / Discard

## Stage II: Analyseblitz (Half-day intensive)

Systematic analysis execution through mock analysis → real analysis → sniff test.

```{r stage2}
# Create analyseblitz template
create_analyseblitz_template("P4.1", analysis_type = "regression")

# Check Stage II progress
stage2_progress_check("P4.1")
```

**Stage II Workflow**:

1. **Prep Phase (15-30 minutes)**
   - Reference gathering
   - Package identification
   - Method clarification

2. **Mock Analysis**
   - Data simulation matching expected structure
   - Complete analysis pipeline on simulated data
   - Debug and refine approach

3. **Real Analysis**
   - Copy-paste mock analysis code
   - Replace simulated data with real data
   - Execute full analysis

4. **Sniff Test**
   - Results plausibility check
   - Effect size reasonableness
   - Direction and magnitude assessment

## Stage III: Draftschreiben (Paper writing)

Complete paper drafting with structured sections and word count targets.

```{r stage3}
# Create draft template
create_draftschreiben_template("P4.1", 
                              paper_type = "journal",
                              target_journal = "Journal of Educational Psychology")

# Track source collection
track_sources(search_strategy = "systematic review + pearl growing", 
             new_sources = 12, 
             project_code = "P4.1")

# Pearl growing reference search
pearl_growing_tracker(seed_paper = "Smith et al. 2023", 
                     direction = "downstream", 
                     found_count = 8)

# Check Stage III progress  
stage3_progress_check("P4.1")
```

**Stage III Features**:
- Word count targets by paper type
- Integration with Stage II results
- Bibliography management
- 80% completion threshold tracking
- Automatic reference to previous stage work

---

# Part III: Analysis Workflow Orchestration

PRoMan's master script system orchestrates complex analyses through modular, reusable components.

## Master Script System

```{r master_scripts}
# Create master analysis script with components
create_master_script("P4.1", 
                    analysis_components = c("data_prep", 
                                          "comprehensive_cleaning",
                                          "descriptive_stats", 
                                          "main_analysis", 
                                          "visualization"),
                    file_sets = list(
                      "state_data" = c("data_AL.csv", "data_CA.csv", "data_TX.csv"),
                      "summary_files" = c("summary_2021.csv", "summary_2022.csv")
                    ))

# Execute complete analysis pipeline
run_master_analysis("P4.1")
```

## Automatic File Set Configuration

```{r auto_file_sets}
# PRoMan can detect your file sets automatically
create_master_script("P4.1")  # Auto-detects file patterns

# Or specify custom file sets
create_master_script("P4.1",
                    file_sets = list(
                      baseline = c("baseline_CA.csv", "baseline_NY.csv"),
                      followup = c("followup_CA.csv", "followup_NY.csv")
                    ))
```

**Master Script Features**:
- **Automatic path configuration** using PRoMan project structure
- **File set integration** for batch processing
- **Component timing** and progress tracking
- **Error handling** with component-level isolation
- **Logging integration** with Blitzschreiben system

## Comprehensive Data Cleaning

The comprehensive cleaning component now includes:

```{r comprehensive_cleaning}
# Automatically generated comprehensive_cleaning component includes:
# - Data quality assessment with reports
# - Suspicious value detection
# - Missing data pattern analysis  
# - ID column detection
# - Variable name standardization
# - Demographic variable cleaning
# - Data dictionary generation
# - Before/after comparison reports
# - All reports saved to personal_notes folder
```

## Component Templates

PRoMan provides intelligent templates for common analysis components:

### Data Preparation Component

```{r data_prep_component}
# Generated template includes:
# - File set loading using PRoMan functions
# - Data cleaning with PRoMan string utilities
# - Standardized output to data/cleaned folder
```

### Descriptive Statistics Component

```{r desc_component}
# Generated template includes:
# - data_snapshot() integration
# - Group-wise summaries
# - Automatic output to personal_notes folder
```

### Main Analysis Component

```{r analysis_component}
# Generated template includes:
# - Model specification area
# - Results capture and export
# - Integration with Stage II mock analysis
```

### Visualization Component

```{r viz_component}
# Generated template includes:
# - Plot creation from analysis results
# - Standardized figure saving
# - Publication-ready formatting
```


## Unified Template Creation

These functions can be used for managing projects across different stages, or jumping in where needed.

```{r unified_templates}
# Create any stage template with one function
create_blitz_template(1, "P4.1", title = "My Research Idea")     # Stage I
create_blitz_template(2, "P4.1", analysis_type = "regression")   # Stage II  
create_blitz_template(3, "P4.1", paper_type = "journal")        # Stage III

# Or use stage names
create_blitz_template("triage", "P4.1")
create_blitz_template("analyse", "P4.1")
create_blitz_template("draft", "P4.1")

# Initialize new project (creates tracker + Stage I template)
init_blitz_project("P4.1", "Research idea title")
```

---

# Part IV: Publication Support

## Academic Figure Management

PRoMan provides comprehensive figure management with automatic registration and metadata tracking:

```{r figure_management}
# Save figure with automatic registration
save_figure(my_plot, 
           figure_name = "main_results_regression",
           figure_type = "main",  # or "supplementary", "exploratory"
           caption = "Regression results showing relationship between X and Y",
           notes = "Used robust standard errors",
           formats = c("png", "pdf"),
           dpi = 300,
           width = 8, 
           height = 6)

# List all registered figures
list_figures(type = "main")

# Update figure metadata without regenerating
update_figure_info("main_results_regression",
                  caption = "Updated caption with more detail",
                  figure_type = "supplementary")

# Create figure summary document
create_figure_summary(output_format = "html")

# Export submission-ready figure package
export_figure_package(figures = "all", 
                     format = "tiff",  # Journal requirement
                     dpi = 600,
                     create_list = TRUE)
```

## Academic Formatting and Styling

```{r academic_formatting}
# Use publication-ready theme
library(ggplot2)
p <- ggplot(data, aes(x, y)) +
  geom_point() +
  theme_academic(base_size = 11, grid = TRUE)

# Colorblind-friendly palettes
colors <- academic_palette("categorical", n = 4)
# Uses Wong (2011) colorblind-safe colors

# Format numbers for publication
p_values <- format_academic(c(0.001, 0.043, 0.52), type = "p_value")
# Returns: "p < .001", "p = .04", "p = .52"

correlations <- format_academic(c(0.234, -0.891), type = "correlation")
# Returns: ".23", "-.89"

# Create matching table and figure
results <- create_table_figure_pair(data,
                                  table_vars = c("outcome", "predictor1", "predictor2"),
                                  plot_type = "bar",
                                  group_var = "condition",
                                  save_pair = TRUE,
                                  pair_name = "descriptive_stats")
```

---

# Part V: Collaboration and Communication

## Collaborator Onboarding

PRoMan can generate comprehensive onboarding packages for new collaborators:

```{r collaboration}
# Create onboarding package
create_collaborator_onboarding(project_code = "P4.1",
                             collaborator_name = "Dr. Smith",
                             package_type = "mixed",  # executive, tutorial, operational, mixed
                             focus_areas = "all")

# This creates:
# - Executive summary (strategic overview)
# - Tutorial walkthrough (methods explanation)
# - Operational details (technical implementation)
# - Navigation guide

# The package automatically extracts information from:
# - Stage I triage cards
# - Stage II analyseblitz files
# - Stage III drafts
# - Data dictionaries
# - Figure registries
```

## Communication Templates

```{r communication}
# Generate email update templates
create_update_template(project_code = "P4.1",
                     update_type = "results")  # progress, milestone, request, results

# Create meeting agendas
create_meeting_agenda(project_code = "P4.1",
                    meeting_type = "progress",  # kickoff, progress, results, planning
                    attendees = c("Dr. Smith", "Dr. Jones"))
```

---

# Part VI: Intelligent Activity Tracking

## Multi-Project Session Management

PRoMan automatically tracks all projects accessed during an R session and prompts for comprehensive logging at exit.

```{r session_tracking}
# PRoMan automatically tracks when you access projects via get_project_paths()
# No manual intervention required!

# Check current pipeline status
blitzschreiben_status()

# See today's recommended tasks
todays_blitz_tasks()

# Get weekly pipeline report
weekly_blitz_report()
```

## Project Logging

```{r enhanced_logging}
# Generate comprehensive project log
generate_project_log(days_back = 7, save_to_file = TRUE)

# Quick status check
quick_project_status()

# The log includes:
# - Recent file activity with emojis for quick scanning
# - Project structure visualization
# - Automatic file set detection
# - Activity inference (e.g., "📊 Data pipeline active")
# - Smart insights about workflow patterns
```

## Smart Exit Logging

When you quit R, PRoMan automatically:

1. **Detects all projects** worked on during the session
2. **Shows session summary** with duration and project list
3. **Prompts for project-specific comments** about next steps/ideas
4. **Generates individual logs** for each project with your comments
5. **Preserves continuity** across work sessions

PRoMan adapts to your environment:
- **In RStudio**: Uses console input for compatibility
- **In R GUI**: Uses graphical dialogs
- **Automatic fallback**: If GUI fails, switches to console

The system now tracks actual project paths during your session, ensuring logs are always generated in the correct locations even when working across multiple project directories.


## Pipeline Management

```{r pipeline_management}
# View complete pipeline status
blitzschreiben_status()
# Shows all projects across stages with recommendations

# Advance projects through stages
advance_to_next_stage("P4.1", 
                     decision = "ready", 
                     next_steps = "Begin mock analysis with simulated data")

# Check what you should work on today
todays_blitz_tasks()
# Provides day-specific recommendations based on Blitzschreiben methodology
```

---

# Complete Workflow Example

## Scenario: New Research Project

```{r complete_workflow}
# 1. Monday: Create new project and do Stage I triage
create_blitz_project("Sleep Quality and Academic Performance", "P5.2")
start_triage_timer(15, "P5.2")
# [Complete triage card]
end_triage_timer("ready_to_share", "Find sleep survey datasets")

# 2. Advance to Stage II
advance_to_next_stage("P5.2", "ready")

# 3. Tuesday: Create analyseblitz template
create_analyseblitz_template("P5.2", analysis_type = "regression")
# [Complete mock analysis in Quarto document]

# 4. Wednesday: Execute real analysis with master script
create_master_script("P5.2", 
                    analysis_components = c("data_prep", "comprehensive_cleaning",
                                          "sleep_analysis", "visualization"))
run_master_analysis("P5.2")

# 5. Thursday: Advance to Stage III and begin writing  
advance_to_next_stage("P5.2")
create_draftschreiben_template("P5.2", paper_type = "journal")

# 6. Friday: Review progress and plan next steps
stage3_progress_check("P5.2")
blitzschreiben_status()  # Overall pipeline health check
```

## Session Exit Example

When you quit R after working on multiple projects:

```
=== R SESSION ENDING ===
Projects worked on this session:

    P5.2
    A3.1
    P2.4

Generate project logs with comments? (y/n): y

--- PROJECT: P5.2 ---
Next steps/ideas: Complete sleep quality variable coding, run sensitivity analysis
✅ Log generated for P5.2

--- PROJECT: A3.1 ---

Next steps/ideas: Submit to conference, prepare presentation slides
✅ Log generated for A3.1

--- PROJECT: P2.4 ---
Next steps/ideas: Address reviewer comments on methods section

✅ Log generated for P2.4

🎉 Multi-project logging completed!

```


---

## File Naming for Optimal Set Detection

- Use consistent suffixes: `data_CA.csv`, `data_NY.csv`
- Include years: `results_2021.csv`, `results_2022.csv`
- Use underscores as separators
- Avoid spaces in filenames
- Include processing stage: `data_CA_raw.csv`, `data_CA_clean.csv`


# Best Practices and Tips

## Blitzschreiben Success Strategies

1. **Respect time limits** - Stage I triage should genuinely take 12-15 minutes
2. **Use mock analysis** - Don't skip the simulation step in Stage II
3. **Maintain pipeline balance** - Don't let any stage get overloaded
4. **Trust the process** - The methodology is designed for momentum, not perfection

## File Organization Tips

- **Use consistent naming** for file sets (state_XX, year_YYYY patterns)
- **Archive regularly** using PRoMan's archive functions
- **Leverage file set detection** - let PRoMan identify patterns automatically
- **Keep Stage II and Stage III connected** - use source() to import results

## Data Quality Best Practices

- Run `data_snapshot()` on new datasets immediately
- Use `flag_suspicious_values()` to catch common missing data codes
- Generate data dictionaries early and update them often
- Use the comprehensive_cleaning component for systematic quality checks

## Figure Management Tips

- Register all figures with `save_figure()` for tracking
- Use descriptive figure names without spaces
- Include both PNG (for viewing) and PDF/TIFF (for publication)
- Update captions and notes as interpretations evolve
- Export figure packages when submitting papers

## Collaboration Strategies

- Generate onboarding packages when adding team members
- Use "executive" package type for PIs and senior collaborators
- Use "operational" package type for RAs doing hands-on work
- Update packages as project evolves
- Use communication templates for consistent updates

## ADHD-Friendly Features

- **Visual pipeline status** - always see where you are
- **Time boundaries** - clear start/stop points prevent hyperfocus
- **Automatic logging** - capture progress without remembering to document
- **Daily task guidance** - know what to work on when
- **Multi-project continuity** - never lose track of what you were doing

## Troubleshooting

**"No Blitzschreiben tracker found"**
- Create your first Blitzschreiben project with `create_blitz_project()`

**Master script errors**
- Check that all component scripts exist in r_programs folder
- Verify file paths using `get_project_paths()`

**Exit logging not working**
- Ensure you loaded PRoMan with `library(PRoMan)`
- Make sure you're in a valid project directory when working

**Pipeline recommendations seem off**
- Run `blitzschreiben_status()` to see current project distribution
- Use `advance_to_next_stage()` to move completed projects forward

**Figure registry issues**
- Make sure to use `save_figure()` instead of direct `ggsave()`
- Check that figures folder exists in personal_notes

**Data cleaning errors**
- Verify column names exist before using cleaning functions
- Check data types - some functions expect numeric columns

**Onboarding package missing information**
- Ensure Blitzschreiben stages have been completed
- Manually edit generated documents to add missing details

**"Can't find project root"**
- Ensure you have the required folders: data/, r_programs/, personal_notes/
- PRoMan searches up to 10 levels from current directory

**"File sets not showing"**
- File sets need at least 2 files with common patterns
- Use `analyze_file_sets()` to see what PRoMan detects
- Check both data/source and data/cleaned folders

**"Exit logging dialog not appearing"**
- In RStudio, check the console for prompts
- The system uses console input in RStudio for better compatibility

**"Master script components missing"**
- Component scripts are auto-generated as templates
- Check r_programs/ folder for component files
- Each component needs format: componentname_PROJECTCODE.R


---

# Advanced Features

## Custom Analysis Components

You can create custom component templates for specialized analyses:

```{r custom_components}
# Create master script with custom components
create_master_script("P6.1",
                    analysis_components = c("data_prep", "survival_analysis", 
                                          "propensity_matching", "sensitivity_tests"))
```

## Integration with External Tools

PRoMan works seamlessly with:
- **Quarto/RMarkdown** for reproducible analysis
- **Reference managers** through .bib file support
- **External data sources** via flexible file loading

## Scaling Up

For large research programs:
- Use project naming conventions (P1.1, P1.2 for related projects)
- Maintain separate Blitzschreiben trackers for different research lines
- Leverage file set detection for multi-site or multi-year studies

---

## Getting Help

- Use `?function_name` for detailed function help
- Check `blitzschreiben_status()` when unsure about pipeline state
- Review project logs in `personal_notes/logs/` for session continuity
- Run `todays_blitz_tasks()` when you need direction
- Check `list_figures()` to see all registered figures
- Review data quality reports in `personal_notes/`
- Generated onboarding packages are starting points - customize them!

## Interactive Reference Guide

```{r help_system}
# Open interactive HTML reference in browser
proman_help()  # Complete reference guide

# Filter by category
proman_help("data")      # Just data management functions
proman_help("blitz")     # Blitzschreiben functions
proman_help("figures")   # Figure management

# Compact mode for side-by-side viewing
proman_help(compact = TRUE)

```


